calibration

### 2017-04-05 Rmd/calibration.Rmd
Initial steps in revision of analysis, including updated calibration and main data sets (addition of new effective extent variable, change of variable names:
sampled_area -> act_ext
eff_ext
study_duration -> act_dur
eff_dur
samp_duration -> samp_dur

Analyses revised through calibration set and data merging of main analysis

```{r, warning = FALSE}
# external fixes and recodes, bring back in
# cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
# cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationffff.xlsx")))
names(cal)
cal[, c("notes", "V30") := NULL]
for(j in names(cal)[c(4:5, 10:24)]) {
  set(cal, j = j, value = as.numeric(cal[[j]]))
}
# length(which(is.na(cal$feature)))

@@

## Variance in extracted statistics
### By DOI
```{r}
varnms <- c("plot_res", "n_sites", "sampled_area", "t_btwn_samp", 
            "study_duration")
mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]  # mean variable val per paper per observer
# variance and mean per paper across observers
vars <- mus[, lapply(.SD, sd, na.rm = TRUE), by = DOI, .SDcols = varnms] 
musmu <- mus[, lapply(.SD, mean, na.rm = TRUE), by = DOI, .SDcols = varnms]

par(mfrow = c(3, 2), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
for(i in varnms) {
  mu <- musmu[, get(i)]
  sdev <- vars[, get(i)]
#   print(paste(i, ":", sdev))
  print(paste(i, ":", round(mu, 2), ":", round(sdev, 2), ":", 
              round(sdev / mu, 2)))
  # plot(mu, sdev / mu, xlab = "mean", ylab = "CV", main = i)
  plot(mu, sdev, xlab = "mean", ylab = "sd", main = i)
}

@@

### By feature (individual records) across DOIs
```{r, results='hold'}
CV <- function(x) sd(x) / mean(x)
rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
          "sampled_area", "t_btwn_samp", "study_duration")
calr <- cal[, rnms, with = FALSE]
# calr[feature != 99, ]
# calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
#      .SDcols = rnms[4:7]]
# calr[feature != 99, .N, by = .(DOI, feature)]
calr[feature != 99, lapply(.SD, CV), by = .(DOI, feature), 
     .SDcols = rnms[4:8]][, lapply(.SD, mean, na.rm = TRUE), 
                          .SDcols = rnms[4:8]]

@@

Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area. [DONE]

Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.  [DONE]

For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts [Chang, Ahmed remaining]

>>>>

```{r, warning = FALSE}
# external fixes and recodes, bring back in
# cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
# cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationfffff.xlsx")))
names(cal)

cal[, c("notes", "V31") := NULL]
nuvars <- c("inout", "feature", "study_year", "st", "plot_res", "n_sites",
            "act_ext", "eff_ext", "samp_dur", "t_btwn_samp", "act_dur", 
            "eff_dur", paste0("comp", 1:3), paste0("struc", 1:3), 
            paste0("func", 1:3))
for(j in nuvars) {
  set(cal, j = j, value = as.numeric(cal[[j]]))
}
# length(which(is.na(cal$feature)))

@@

## Variance in extracted statistics
### By DOI
```{r}
varnms <- c("plot_res", "n_sites", "act_ext", "eff_ext", "t_btwn_samp", 
            "act_dur", "eff_dur")
mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]  # mean variable val per paper per observer
# variance and mean per paper across observers
vars <- mus[, lapply(.SD, sd, na.rm = TRUE), by = DOI, .SDcols = varnms] 
musmu <- mus[, lapply(.SD, mean, na.rm = TRUE), by = DOI, .SDcols = varnms]

par(mfrow = c(3, 3), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
for(i in varnms) {
  mu <- musmu[, get(i)]
  sdev <- vars[, get(i)]
#   print(paste(i, ":", sdev))
  print(paste(i, ":", round(mu, 2), ":", round(sdev, 2), ":", 
              round(sdev / mu, 2)))
  # plot(mu, sdev / mu, xlab = "mean", ylab = "CV", main = i)
  plot(mu, sdev, xlab = "mean", ylab = "sd", main = i)
}

@@

### By feature (individual records) across DOIs
```{r, results='hold'}
CV <- function(x) sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
          "act_ext", "eff_ext", "t_btwn_samp", "act_dur", "eff_dur")
calr <- cal[, rnms, with = FALSE]

# calr[feature != 99, ]
# calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
#      .SDcols = rnms[4:7]]
# calr[feature != 99, .N, by = .(DOI, feature)]
calr[feature != 99, lapply(.SD, CV), by = .(DOI, feature), 
     .SDcols = rnms[4:10]][, lapply(.SD, mean, na.rm = TRUE), 
                          .SDcols = rnms[4:10]]

@@
